---
title: "Exploring Seattle's Unreinforced Buildings"
author: "bxie"
date: "11/4/2019"
---

## Unreinforced Masonry Buildings in Seattle
The City of Seattle has identified many buildings as in need of earthquake
retrofitting. Let's see where they are and how bad it is!

Data from [data.seattle.gov](https://data.seattle.gov/Permitting/Unreinforced-Masonry-Buildings/54qs-2h7f)

For more info on ggplot2, see [ggplot quick ref](http://sape.inf.usi.ch/quick-reference/ggplot2/geom)

```{r setup, include=FALSE, eval=FALSE}
library(dplyr)
library(ggplot2)
library(leaflet)
library(stringr)
```

Reading in data
```{r cars}
df <- read.csv("data/seattle_buildings.csv", stringsAsFactors = FALSE)
View(df)
```


```{r}
# simplifying risk categorization
df <- df %>% mutate(risk = str_split(Preliminary.Risk.Category, " ")  %>% 
                  as.data.frame() %>% 
                  t %>% 
                  data.frame(stringsAsFactors = F) %>% 
                  pull(1))
```

Risk by neighborhood
```{r}
# TODO: In df_neighborhood, group by risk and Neighborhood and then summarize
# the number of observerations in each group w/ n()
df_neighborhood <- df %>% ...


# TODO: add labels to the plot with labs() and ylab()
ggplot(df_neighborhood, aes(fill=risk, y = n, x = Neighborhood)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5)) # rotate x axis labels

# TODO: Now add 'position = "fill"' as a parameter in geom_bar. 
# How does the plot change? What changes about the information conveyed?


```

Adding columns for map visualizations
```{r}
# TODO: View df and look at df$Long.Lat. How would you 
# extract the information you need for map visualizations

# Getting longitude and latitude from Long.Lat column
lng_vec <- lapply(df$Long.Lat, function(x){
  as.numeric(str_split(substr(x, 8, str_length(x)-1), " ")[[1]][1])
})
lat_vec <- lapply(df$Long.Lat, function(x){
  as.numeric(str_split(substr(x, 8, str_length(x)-1), " ")[[1]][2])
})
df$lng <- unlist(lng_vec)
df$lat <- unlist(lat_vec)

```


```{r}
# TODO: create column "info" which has the address, a breakline ("<br>"), 
# and risk level
df <- df %>% mutate(...)

# TODO: Starting with the leaflet of the df beflow, add tiles, and add markers specifying
# longitude (lng), latitude (lng), and popup (info). Don't forget the tilde!
leaflet(df) %>% ...
```



--------------------------- (time permitting) ---------------------------

# Fix these bad ggplots!
```{r}
if (!require(data.table)) { install.packages('data.table'); require(data.table) }
if (!require(plyr)) { install.packages('plyr'); require(plyr) }
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars, aes(x=cyl)) + geom_bar(fill='darkgoldenrod2') +
  theme(panel.background = element_rect(fill = 'steelblue'),
  panel.grid.major = element_line(colour = "firebrick", size=3),
  panel.grid.minor = element_line(colour = "blue", size=1))
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
freq_cars_by_cyl <- as.data.frame(table(mtcars %>% select(cyl))) # freq table of # cars by cylinder
colnames(freq_cars_by_cyl)[colnames(freq_cars_by_cyl) == "Var1"] <- "cyl" # rename col

ggplot(as.data.frame(freq_cars_by_cyl),
       aes(x = "", y = Freq, fill=cyl)) +
  geom_bar(width=1, stat = 'identity')  + # create bar plot
  coord_polar(theta="y") # turn to pie chart
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars, aes(x=disp, y=hp)) + 
    geom_point(
        color="pink",
        fill="blue",
        shape=21,
        alpha=0.5,
        size=50,
        stroke = 4
        )
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
```{r}
ggplot(mtcars) +
  geom_point(aes(wt, hp, color=cyl)) +
  scale_colour_gradient(low = "red", high="green")
```

# TODO: With a partner, identify at least 2 issues for this plot. Then fix them!
_hint_: check out facet_grid
```{r}
head(iris)
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + geom_point(shape=4, size=20)
```

